---
title: "GLMM"
output:
  ioslides_presentation:
    incremental: true
    widescreen: true
    smaller: true
    transition: faster
---

<style>
/* This targets the slide transition speed (default 'faster' is ~0.2s) */
slides > slide {
  -webkit-transition: all 0.1s ease-in-out;
  -moz-transition: all 0.1s ease-in-out;
  -o-transition: all 0.1s ease-in-out;
  transition: all 0.1s ease-in-out;
}
.scrollable-slide {
  overflow-y: auto !important;
  height: 500px;
}
.hidden-slide {
  display: none;
}
</style>


```{r setup, include=FALSE} 
knitr::opts_chunk$set(
  echo = FALSE, 
  comment = NA, 
  message = FALSE, 
  warning = FALSE,
  fig.align = 'center',
  fig.height = 4,
  fig.width = 6)
options(scipen = 1, digits = 5)
```

```{r}
depression_data <- read.table("Depression.dat", header = TRUE)
```


```{r}
library(mlmRev)
data(Contraception)
library(kableExtra)
library(dplyr)
library(tidyr)
library(MASS)
library(pROC)
library(caret)
library(knitr)
```

# GLM Limitations

## GLM

- Requires independent samples

## Repeated measures {.build}

- The data shows multiple observations per individual (Case ID). Each individual has data on severity, drug, and outcome measurement in three different time points. (decription on page 456)

```{r}
library(DT)

subset_data <- depression_data %>% 
  filter(case %in% c(21, 22, 23))

datatable(subset_data, options = list(pageLength = 9,
                                      searching = FALSE, 
                                      lengthChange = FALSE, 
                                      dom = 't'))
```

- GLM would assume each row is independent from others, as if we have 9 independent samples.
- Observations within clusters are usually positively correlated, tending to be more alike than observations from different clusters

## Repeated measures {.build}

Reshaping highlights that we only have 3 independent samples, each with multiple observations.


```{r}
subset_data1 <- depression_data %>% filter(case %in% c(21, 22, 23)) %>% pivot_wider(names_from = 'time', values_from = 'outcome')

datatable(subset_data1, options = list(pageLength = 9,
                                      searching = FALSE, 
                                      lengthChange = FALSE, 
                                      dom = 't'))
```

## Natural groupings in data {.build}

- This shows yes-no responses on use of contraceptives among women of different ages from three out of 60 districts in Bangladesh.

```{r, fig.width=6, fig.height=3}
cont_filtered1 <- Contraception %>% filter(district %in% c(1, 34, 30))#c(34, 21, 61))#c(18, 30, 46)) 
cont_filtered2 <- cont_filtered1
cont_filtered2$district <- "combined"
rbind(cont_filtered1, cont_filtered2)%>% 
  ggplot(aes(x = district, y = age, color = use))+
  facet_grid(.~district, scales = "free_x") +
  geom_boxplot()  +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed")+ theme_bw() +
    theme(
    axis.text.x = element_blank(),  # Removes the labels (e.g., "18", "30")
    axis.ticks.x = element_blank(), # Removes the small tick marks
    axis.title.x = element_blank()  # Optional: Removes the "district" title
  )
  # group_by(district, use) %>% count()
```
- Ordinary analyses, such as GLM, that ignore the correlation and treat within-cluster observations the same as between-cluster observations produce invalid standard errors, tending to be too small for between-cluster effect estimates and too large for within-cluster effects.


## Natural groupings in data {.build}

```{r}
num_district <- Contraception %>% pull(district) %>% unique() %>% length()
```

- Why not use districts as one of the explanatory variables?
  - This dataset actually includes `r num_district` districts. Some of which have very few observations.
  
```{r, fig.width = 8, fig.height = 4}
Contraception %>% group_by(district) %>% count() %>% arrange(n) %>%
  ggplot(aes(x = reorder(district, desc(n)), y = n)) +
  geom_col() + theme_bw()  +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  #geom_hline(aes(yintercept = min(n)), color = "red", linetype = "dashed") +
  labs(title = 'Number of observations per district')
```

```{r}
dist <- Contraception %>% 
  group_by(district) %>% 
  count() %>% 
  arrange(desc(n)) %>% 
  head(10) %>% 
  pull(district)
```

## GLM

- Requires independent samples
- Observations within clusters are usually positively correlated, tending to be more alike than observations from different clusters.  
- Ordinary analyses that ignore the correlation and treat within-cluster observations the same as between-cluster observations produce invalid standard errors, tending to be too small for between-cluster effect estimates and too large for within-cluster effects.
- Solution: add cluster-level terms to the model that take the same value for each observation in a cluster. They are unobserved and when treated as varying randomly among clusters, they are called random effects.

# Random Effects Model

```{r}
# Contraception %>% #filter(district %in% c(18, 30, 46)) %>% 
#   ggplot(aes(x = factor(fake_district), y = age, color = use))+
#   facet_grid(.~fake_district, scales = "free_x") +
#   geom_boxplot()  +
#   geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed")+ theme_bw()
```



## Definitions {.build}

- **Fixed effects** are parameters that describe a factor's effects in an ordinary GLM. They apply to all categories of interest, such as gender, treatment, or age groupings. ss
- **Random effects** usually apply to a sample. For a study that makes repeated observations on a sample of subjects, the model treats observations from a given subject as a cluster, and it has a random effect for each of the subject.
- GLMs extend ordinary regression by allowing nonnormal responses and a link function of the mean. The GLMM is a further extension that permits random effects as well as fixed effects in a linear predictor.

## GLMM {.build}

- you have $I$ clusters: $i = 1, \dots, I$ with $T_i$ observations in the $i$th cluster
- $y_{it}$: observation $t$ in cluster $i$
- the number of observations may vary by cluster
- $\mathbf{x}_{it}$: column vector of values of explanatory variables for $y_{it}$
- $\mathbf{u}_i$: vector of random effects for cluster $i$. This is often univariate. This is assumed to have a multivariate normal distribution $\mathcal{N}(\mathbf{0}, \boldsymbol{\Sigma})$. The covariance matrix $\boldsymbol{\Sigma}$ depends on unknown variance components and possible also correlation parameters.
- Conditional on $\mathbf{u}_i$, a GLMM resembles an ordinary GLM. Conditional on $\mathbf{u}_i$, standard model fitting treats $\{y_{it}\}$ as independent over $t$ and $i$.
- Let $\mu_{it} = E(Y_{it} \mid \mathbf{u}_i)$. The linear predictor for a GLMM for link function $g(\cdot)$ and fixed effect model parameters $\boldsymbol{\beta}$ has the form

$$
g(\mu_{it}) = \mathbf{x}^T_{it}\boldsymbol{\beta} + \mathbf{z}_{it}^T \mathbf{u}_i
$$



## GLMM - Random Intercept {.build .scrollable-slide}

Let $T = 2$, and $\boldsymbol{\beta} = \begin{bmatrix} \alpha \\ \beta\end{bmatrix}$.

For a single observation $(i, t)$, the formula is

$$
g(\mu_{it}) = \mathbf{x}^T_{it}\boldsymbol{\beta} + \mathbf{z}_{it}^T \mathbf{u}_i
$$

If $\mathbf{u}_i$ is univariate, we can either have a random slope or a random intercept. Suppose we have a random intercept. Note that $u_i \sim\mathcal{N}(0, \sigma^2)$

$$
\begin{aligned}
\begin{bmatrix} 
g(\mu_{11}) \\ 
g(\mu_{12}) \\ 
g(\mu_{21}) \\ 
g(\mu_{22}) \\ 
\vdots \\
g(\mu_{I1}) \\ 
g(\mu_{I2}) \\ 
\end{bmatrix} 
&= 
\begin{bmatrix} 
\mathbf{x}_{11}^\top \\ 
\mathbf{x}_{12}^\top \\ 
\mathbf{x}_{21}^\top \\ 
\mathbf{x}_{22}^\top \\ 
\vdots \\
\mathbf{x}_{I1}^\top \\ 
\mathbf{x}_{I2}^\top \\ 
\end{bmatrix}
\begin{bmatrix} \alpha \\ \beta\end{bmatrix}
+ 
\begin{bmatrix} 
\mathbf{z}_{11}^\top \\ 
\mathbf{z}_{12}^\top \\ 
\mathbf{z}_{21}^\top \\ 
\mathbf{z}_{22}^\top \\ 
\vdots \\
\mathbf{z}_{I1}^\top \\ 
\mathbf{z}_{I2}^\top \\ 
\end{bmatrix}
\begin{bmatrix} u_1 \\ u_2 \\ \vdots \\ u_I \end{bmatrix} \\
& = 
\begin{bmatrix} 
1 & x_{11} \\ 
1 & x_{12} \\ 
1 & x_{21} \\ 
1 & x_{22} \\ 
\vdots \\
1 & x_{I1} \\ 
1 & x_{I2} \\ 
\end{bmatrix}
\begin{bmatrix} \alpha \\ \beta\end{bmatrix} 
+ 
\begin{bmatrix} 
1 & 0 & \cdots & 0 \\ 
1 & 0 & \cdots & 0 \\ 
0 & 1 & \cdots & 0 \\ 
0 & 1 & \cdots & 0 \\ 
\vdots & \vdots & \ddots & \vdots \\
0 & 0 & \cdots & 1 \\ 
0 & 0 & \cdots & 1 \\ 
\end{bmatrix}
\begin{bmatrix} u_1 \\ u_2 \\ \vdots \\ u_I \end{bmatrix} \\
g[E(\mathbf{Y} \mid \mathbf{u})] = g(\boldsymbol{\mu}) 
&= \mathbf{X}_{(2I)\times 2} \boldsymbol\beta_{2 \times 1} 
+ \mathbf{Z}_{(2I)\times I} \mathbf{u}_{I \times 1}, 
\qquad \text{matrix notation}
\\
\begin{bmatrix} 
g(\mu_{11}) \\ 
g(\mu_{12}) \\ 
g(\mu_{21}) \\ 
g(\mu_{22}) \\ 
\vdots \\
g(\mu_{I1}) \\ 
g(\mu_{I2}) \\ 
\end{bmatrix}
&=
\begin{bmatrix} 
(\alpha + u_1) + x_{11}\beta \\ 
(\alpha + u_1) + x_{12}\beta \\ 
(\alpha + u_2) + x_{21}\beta \\ 
(\alpha + u_2) + x_{22}\beta \\ 
\vdots \\
(\alpha + u_I) + x_{I1}\beta \\ 
(\alpha + u_I) + x_{I2}\beta \\ 
\end{bmatrix}
\end{aligned}
$$

- Link(Average outcome) = General trend (Fixed) + Cluster deviation (Random)
- $\alpha$ and $\beta$ are the fixed effects while $u_1$ and $u_2$ are the random effects
- If $Y_{it} \mid u_i \sim \text{Bernoulli}(\mu_{it})$, $g = \text{logit}$, and $x_{it}$ indicates whether treatment is administered or not, we have a logistic GLMM with random intercept for binary matched pairs.
  - an extension allows $T > 2$ observations in each cluster. The random intercept model has form below and early applications were in psychometrics (item response models).
  
  $$
  \text{logit}[P(Y_{it}=1 \mid u_i)] = \alpha + \beta_t + u_i
  $$
  
    - The item-response model with random intercept is a special case of an important class of random effects model for binary data called logistic-normal models for binary data with the form
    
    $$
    \text{logit}[P(Y_{it}=1 \mid u_i)] = \mathbf{x}^T_{it} \boldsymbol{\beta}+ u_i
    $$

## GLMM - Random Slope {.build .scrollable-slide}

Let $I = 2$, $T = 2$, and $\boldsymbol{\beta} = \begin{bmatrix} \alpha \\ \beta\end{bmatrix}$.

For a single observation $(i, t)$, the formula is

$$
g(\mu_{it}) = \mathbf{x}^T_{it}\boldsymbol{\beta} + \mathbf{z}_{it}^T u_i
$$

If $u_i$ is a **random slope** for the covariate $x_{it}$, then:

$$
\begin{aligned}
\begin{bmatrix} 
g(\mu_{11}) \\ 
g(\mu_{12}) \\ 
g(\mu_{21}) \\ 
g(\mu_{22}) \\ 
\end{bmatrix} &= 
\begin{bmatrix} 
1 & x_{11} \\ 
1 & x_{12} \\ 
1 & x_{21} \\ 
1 & x_{22} \\ 
\end{bmatrix}
\begin{bmatrix} \alpha \\ \beta\end{bmatrix}
+ 
\begin{bmatrix} 
x_{11} & 0 \\ 
x_{12} & 0 \\ 
0 & x_{21} \\ 
0 & x_{22} \\ 
\end{bmatrix}
\begin{bmatrix} u_1 \\ u_2\end{bmatrix} \\
\\
\begin{bmatrix} 
g(\mu_{11}) \\ 
g(\mu_{12}) \\ 
g(\mu_{21}) \\ 
g(\mu_{22}) \\ 
\end{bmatrix}
&=
\begin{bmatrix} 
\alpha + x_{11}(\beta + u_1) \\ 
\alpha + x_{12}(\beta + u_1) \\ 
\alpha + x_{21}(\beta + u_2) \\ 
\alpha + x_{22}(\beta + u_2) \\ 
\end{bmatrix}
\end{aligned}
$$

- The term $(\beta + u_i)$ represents the cluster-specific slope.
- $\alpha$ and $\beta$ are the fixed effects while $u_1$ and $u_2$ are the random slope deviations.

## GLMM - Random Intercept & Slope {.build .scrollable-slide}

Let $I = 2$, $T = 2$, and $\boldsymbol{\beta} = \begin{bmatrix} \alpha \\ \beta\end{bmatrix}$. 

Now, $\mathbf{u}_i = \begin{bmatrix} u_{0i} \\ u_{1i} \end{bmatrix}$ is a **vector** (bivariate) containing a random intercept and a random slope.

- $\mathbf{z}_{it}^T = [1, x_{it}]$

$$
\begin{aligned}
\begin{bmatrix} 
g(\mu_{11}) \\ 
g(\mu_{12}) \\ 
g(\mu_{21}) \\ 
g(\mu_{22}) \\ 
\end{bmatrix} &= 
\begin{bmatrix} 
1 & x_{11} \\ 
1 & x_{12} \\ 
1 & x_{21} \\ 
1 & x_{22} \\ 
\end{bmatrix}
\begin{bmatrix} \alpha \\ \beta\end{bmatrix}
+ 
\begin{bmatrix} 
1 & x_{11} & 0 & 0 \\ 
1 & x_{12} & 0 & 0 \\ 
0 & 0 & 1 & x_{21} \\ 
0 & 0 & 1 & x_{22} \\ 
\end{bmatrix}
\begin{bmatrix} u_{01} \\ u_{11} \\ u_{02} \\ u_{12} \end{bmatrix} \\
\\
\begin{bmatrix} 
g(\mu_{11}) \\ 
g(\mu_{12}) \\ 
g(\mu_{21}) \\ 
g(\mu_{22}) \\ 
\end{bmatrix}
&=
\begin{bmatrix} 
(\alpha + u_{01}) + x_{11}(\beta + u_{11}) \\ 
(\alpha + u_{01}) + x_{12}(\beta + u_{11}) \\ 
(\alpha + u_{02}) + x_{21}(\beta + u_{12}) \\ 
(\alpha + u_{02}) + x_{22}(\beta + u_{12}) \\ 
\end{bmatrix}
\end{aligned}
$$

- **Fixed Part:** $\alpha$ (intercept) and $\beta$ (slope).
- **Random Part:** $u_{0i}$ (intercept deviation) and $u_{1i}$ (slope deviation) for each cluster $i$.
- $\mathbf{Z}$ is now a **block-diagonal matrix** where each block is the $\mathbf{z}_{it}$ values for that cluster.





## Heterogeneity in Logistic-Normal models


Take 2 observations from different clusters. Let them be $(x_{it}, y_{it})$ and $(x_{hs}, y_{hs})$. We look at their difference in log-odds.

$$
\text{logit}[P(Y_{it}=1 \mid u_i)] = \mathbf{x}^T_{it} \boldsymbol{\beta}+ u_i
$$

$$
\text{logit}[P(Y_{hs}=1 \mid u_h)] = \mathbf{x}^T_{hs} \boldsymbol{\beta}+ u_h
$$


$$
\begin{aligned}
\text{logit}[P(Y_{it}=1 \mid u_i)] - \text{logit}[P(Y_{hs}=1 \mid u_h)] &=
\mathbf{x}^T_{it} \boldsymbol{\beta} + u_i - \mathbf{x}^T_{hs} \boldsymbol{\beta} - u_h \\
&= (\mathbf{x}^T_{it}  - \mathbf{x}^T_{hs}) \boldsymbol{\beta} + (u_i - u_h) \\
&= (\mathbf{x}_{it}  - \mathbf{x}_{hs})^T \boldsymbol{\beta} + (u_i - u_h)
\end{aligned}s
$$


- Fixed part: $(\mathbf{x}_{it}  - \mathbf{x}_{hs})^T \boldsymbol{\beta}$
- Random part: $(u_i - u_h) \sim \mathcal{N}(0, 2 \sigma^2)$
- $u_i - u_h$ captures additional variation in log-odds due to cluster-level heterogeneity.

If $\sigma = 0$, there is no variation across clusters - the logistic-normal model simplifies to the ordinary logistic regression model, treating all observations as independent.

$$
\begin{aligned}
\text{logit}[P(Y_{it}=1 \mid u_i)] - \text{logit}[P(Y_{hs}=1 \mid u_h)]
&= (\mathbf{x}_{it}  - \mathbf{x}_{hs})^T \boldsymbol{\beta} + (u_i - u_h) \\
&= (\mathbf{x}_{it} - \mathbf{x}_{hs})^T \boldsymbol{\beta}
\end{aligned}
$$

If $\sigma > 0$, $(\mathbf{x}_{it}  - \mathbf{x}_{hs})^T \boldsymbol{\beta} + (u_i - u_h)$ only reduces to $(\mathbf{x}_{it} - \mathbf{x}_{hs})^T \boldsymbol{\beta}$ if $u_i = u_h$ --- when observations come from the same cluster.

## Form: GLM vs GLMM {.big-text .build}

Suppose we use the logit link:

GLM (marginal model)

$$
\text{logit}[P(Y_{it}=1)] = \mathbf{x}^T_{it} \boldsymbol{\beta}
$$

GLMM (conditional on random effects); $u_i \sim \mathcal{N}(0,\sigma^2)$

$$
\text{logit}[P(Y_{it}=1 \mid u_i)] = \mathbf{x}^T_{it} \boldsymbol{\beta}+ u_i
$$

\smallskip
\small
Note: The marginal probability $P(Y_{it}=1)$ in GLMM has no closed form.

## Random effects: unmeasured predictors {.build}

The random effects sometimes represent heterogeneity caused by omitting certain explanatory variables. Suppose we only have $p-1$ predictors,

$$
g[E(\mathbf{y})] = g(\boldsymbol{\mu}) = \mathbf{X}_{n \times (p-1)} \boldsymbol{\beta}_{(p-1)\times 1}
$$

But the model could have been better explained in the presence of the $p$th variable which is unmeasured.

The random effects model relate to methods of dealing with unmeasured predictors. The random effects part of the linear predictor reflects terms that would be in the fixed effects part if those explanatory variables had been included.

$$
g[E(\mathbf{y})] = \mathbf{X}_{n \times (p-1)} \boldsymbol{\beta}_{(p-1)\times 1} + \mathbf{Z}_{n \times q}\mathbf{u}_{q \times 1}
$$

## Random effects: random measurement error in the explanatory variables {.build}

Random effects also sometimes represent random measurement error in the explanatory variables. 

If we replace a particular $x_{it}$ by $x^*_{it} + \epsilon_i$, with $x^*_{it}$ the true values and $\epsilon_i$ the measurement error, then $\epsilon_i\beta$ can be absorbed in the random effects term.


$$
\begin{aligned}
g(\mu_{it}) &= \mathbf{x}^T_{it}\boldsymbol{\beta} + \mathbf{z}_{it}^T \mathbf{u}_i
\\
&= (\mathbf{x}^{T*}_{it} + \epsilon_i)\boldsymbol{\beta} + \mathbf{z}_{it}^T \mathbf{u}_i
\\
&= \mathbf{x}^{T*}_{it}\boldsymbol{\beta} + \epsilon_i\boldsymbol{\beta} + \mathbf{z}_{it}^T \mathbf{u}_i
\end{aligned}
$$


## Random effects: explaining overdispersion {.build}

Random effects also provide a mechanism for explaining overdispersion in basic models not having those effects.

- GLM Poisson

$$
Var(Y) = \mu
$$

- GLM Negative Binomial

$$
Var(Y) = \mu + \phi \mu^2, \qquad \phi > 0
$$

- GLMM

$$
Var(Y) = \mu + \mu^2 (\exp ( \sigma^2 ) - 1)
$$


# Example

## Dataset

```{r}
library(mlmRev)
data(Contraception)
head(Contraception)
```

## Fit a GLM {.build .scrollable-slide}

We ignore the district variable.

```{r echo = TRUE}
model_glm  <- glm(use ~ age + urban + livch, 
                  family = binomial, 
                  data = Contraception)
```

```{r}
summary(model_glm)
```


## Fit a GLMM {.build .scrollable-slide}

We use random intercept model.

```{r echo = TRUE}
model_glmm0 <- glmer(use ~ age + urban + livch + (1 | district), 
                    family = binomial, 
                    data = Contraception)
```

```{r}
summary(model_glmm0)
```

## Comparison {.build .scrollable-slide}

```{r fig.height = 5, fig.width = 7}
long_age_range <- seq(-250, 250, length.out = 1201)
districts <- c(1, 30, 34)#c(18, 30, 46)#unique(Contraception$district)
fake_districts <- c(1, 2, 3)
urban <- 'Y'#unique(Contraception$fake_district)
livch <- 1
grid_long <- expand.grid(age = long_age_range, 
                         district = districts,
                         livch = livch,
                         urban = urban)
grid_long <- grid_long %>% mutate(district = factor(district),
                                  livch = factor(livch))

grid_long$prob_glmm <- predict(model_glmm0, newdata = grid_long, type = "response", re.form = NULL)
grid_long$prob_glm  <- predict(model_glm,  newdata = grid_long, type = "response")

ggplot(grid_long, aes(x = age)) +

  
  geom_line(aes(y = prob_glmm, group = district, linetype = district),
            color = "blue", alpha = 0.5) +

  geom_line(aes(y = prob_glm,color = "GLM"),
            color = "red", size = 0.5) +
  scale_linetype_manual(values = c("longdash","solid", "dotted")) + 

  geom_hline(yintercept = 0.5, linetype = "dotted", alpha = 0.5) +
  geom_vline(xintercept = 0, linetype = "dotted", alpha = 0.5) +
  theme_minimal() +
  labs(title = "GLM vs. GLMM",
       subtitle = "Red: GLM | Blue: GLMM Districts",
       x = "Age (Centered/Scaled)",
       y = "Predicted Probability") + theme_bw()
```


## What happens if {.build .scrollable-slide}

... we fit a GLMM to a dataset but the clusters do not vary? 

It gets reduced to a GLM. 

```{r}
set.seed(658) # still diff
set.seed(7)
Contraception <- Contraception[sample(nrow(Contraception)), ]
Contraception$fake_district <- c(rep(1, 645), rep(2, 645), rep(3, 644))
```

### `age` per district

```{r}
Contraception %>% 
  ggplot(aes(x = fake_district, y = age, color = use))+
  facet_grid(.~fake_district, scales = "free_x") +
  geom_boxplot()  +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed")+ theme_bw() +
    theme(
    axis.text.x = element_blank(),  # Removes the labels (e.g., "18", "30")
    axis.ticks.x = element_blank(), # Removes the small tick marks
    axis.title.x = element_blank()  # Optional: Removes the "district" title
  )
```

### `livch` per district

```{r}
Contraception %>% 
  group_by(fake_district, livch) %>% 
  summarise(prop_using = mean(as.character(use) == "Y")) %>%
  ggplot(aes(x = livch, y = prop_using)) +
  geom_col() +
  facet_grid(~fake_district) +
  theme_bw()
```

### `urban` per district

```{r}
Contraception %>% 
  group_by(fake_district, urban) %>% 
  summarise(prop_using = mean(as.character(use) == "Y")) %>%
  ggplot(aes(x = urban, y = prop_using)) +
  geom_col() +
  facet_grid(~fake_district) +
  theme_bw()
```

```{r echo = TRUE}
model_glmm1 <- glmer(use ~ age + urban + livch + (1 | fake_district), 
                    family = binomial, 
                    data = Contraception)
```


```{r}
summary(model_glmm1)
```


```{r fig.height = 5, fig.width = 7}
grid_long <- expand.grid(age = long_age_range, 
                         fake_district = fake_districts,
                         livch = livch,
                         urban = urban)
grid_long <- grid_long %>% mutate(district = factor(fake_district),
                                  livch = factor(livch))

grid_long$prob_glmm <- predict(model_glmm1, newdata = grid_long, type = "response", re.form = NULL)
grid_long$prob_glm  <- predict(model_glm,  newdata = grid_long, type = "response")

ggplot(grid_long, aes(x = age)) +
  geom_line(aes(y = prob_glmm, group = fake_district, linetype = district),
            color = "blue", alpha = 0.25) +

  geom_line(aes(y = prob_glm,color = "GLM"),
            color = "red", alpha = 1, linetype = "dotted") +
  scale_linetype_manual(values = c("longdash","solid", "dotted")) + 

  geom_hline(yintercept = 0.5, linetype = "dotted", alpha = 0.5) +
  geom_vline(xintercept = 0, linetype = "dotted", alpha = 0.5) +
  theme_minimal() +
  labs(title = "GLM vs. GLMM",
       subtitle = "Red: GLM | Blue: GLMM Fake Districts",
       x = "Age (Centered/Scaled)",
       y = "Predicted Probability") + theme_bw()
```





$$
\begin{aligned}
\text{logit}\big(P(\text{use}=1 \mid u_j)\big) &= (\beta_0 + u_j) + \beta_1 \,\text{age} + \beta_2 \,\text{urban} + \beta_3 \,\text{livch}
\\
\text{log} \frac{\big(P(\text{use}=1 \mid u_j)\big)}{1-\big(P(\text{use}=1 \mid u_j)\big)} &= (\beta_0 + u_j) + \beta_1 \,\text{age} + \beta_2 \,\text{urban} + \beta_3 \,\text{livch} 
\\
\frac{\big(P(\text{use}=1 \mid u_j)\big)}{1-\big(P(\text{use}=1 \mid u_j)\big)} &= \exp \left[ (\beta_0 + u_j) + \beta_1 \,\text{age} + \beta_2 \,\text{urban} + \beta_3 \,\text{livch} \right]
\\
&= \exp (\beta_0 + u_j) \times \exp (\beta_1 \,\text{age} + \beta_1) \times \exp (\beta_2 \,\text{urban}) \times \exp (\beta_3 \,\text{livch})
\\
&= \exp (\beta_0 + u_j) \times \exp (\beta_1 \,\text{age}) \times \exp(\beta_1) \times \exp (\beta_2 \,\text{urban}) \times \exp (\beta_3 \,\text{livch})
\end{aligned}
$$


```{r}
summary(model_glm)
```


```{r}
summary(model_glmm0)
```

The model includes a district-level random intercept, indicating that baseline contraceptive use varies across districts. Districts with higher random intercepts have higher baseline probabilities of use, after adjusting for individual-level covariates.

```{r}
# ranef(model_glmm0)$district %>% arrange(`(Intercept)`)
```

```{r}
summary(model_glmm1)
```

```{r}
# set.seed(71123)
set.seed(658) # still diff
set.seed(7)
Contraception <- Contraception[sample(nrow(Contraception)), ]
Contraception$fake_district <- c(rep(1, 645), rep(2, 645), rep(3, 644))



# AIC(model_glm, model_glmm)
# ranef(model_glmm)
```


```{r}
unique(Contraception$livch)
```

```{r}

```



```{r}
library(dplyr)

Contraception <- Contraception %>%
  group_by(district) %>%
  mutate(
    age_mean = mean(age),
    age_within = age - age_mean
  ) %>%
  ungroup()
```

```{r}
# GLM (ignores clustering)
glm_fit <- glm(use ~ age_mean + age_within + urban + livch,
               family = binomial, data = Contraception)

# GLMM (accounts for clustering)
library(lme4)
glmer_fit <- glmer(use ~ age_mean + age_within + urban + livch + (1 | district),
                   family = binomial, data = Contraception)
```

```{r}
summary(glm_fit)
```
```{r}
summary(glmer_fit)
```








```{r}
model_glm  <- glm(outcome ~ severity + drug + time + drug:time , family = binomial, data = depression_data)

model_glmm <- glmer(outcome ~ severity + drug + time + drug:time + (1 | case), family = binomial, data = depression_data)
```



# Appendixs


## Logistic-Normal GLMM {.build .scrollable-slide .hidden-slide} 

Let $I = 2$, $T = 2$, and $\boldsymbol{\beta} = \begin{bmatrix} \alpha \\ \beta\end{bmatrix}$. 
Assume $y_{it} \sim \text{Bernoulli}(\pi_{it})$.

For a single observation $(i, t)$, the model reduces to:

$$
\text{logit}(\pi_{it}) = \alpha + \beta x_{it} + u_i
$$

where $u_i \stackrel{iid}{\sim} N(0, \sigma^2)$.

- **Link Function**: $g(\mu_{it}) = \log\left(\frac{\pi_{it}}{1 - \pi_{it}}\right)$
- **Random Intercept**: $u_i$ represents the cluster-specific deviation from the fixed intercept $\alpha$.

$$
\begin{aligned}
\begin{bmatrix} 
\text{logit}(\pi_{11}) \\ 
\text{logit}(\pi_{12}) \\ 
\text{logit}(\pi_{21}) \\ 
\text{logit}(\pi_{22}) \\ 
\end{bmatrix} & = 
\begin{bmatrix} 
1 & x_{11} \\ 
1 & x_{12} \\ 
1 & x_{21} \\ 
1 & x_{22} \\ 
\end{bmatrix}
\begin{bmatrix} \alpha \\ \beta\end{bmatrix} + 
\begin{bmatrix} 
1 & 0 \\ 
1 & 0 \\ 
0 & 1 \\ 
0 & 1 \\ 
\end{bmatrix}
\begin{bmatrix} u_1 \\ u_2\end{bmatrix} \\
\end{aligned}
$$



## Logistic GLMM with Random intercept for binary matched pairs

(matched pair i induces the cluster, t can be thought of as the treatment)

Cluster $i$ consists of the responses $(y_{i1}, y_{i2})$ for matched pair $i$. Observation $t$ in cluster $i$ has $y_{it} = 1$ (success) or 0 (failure), $t =  1, 2$. Let

$$
\mu_{it} =  P(Y_{it} =1 \mid u_i ) \\
g(\cdot) \text{ be the logit link} \\
\boldsymbol{\beta}^T = (\alpha, \beta) \\
\mathbf{x}^T_{i1} =(1,0) \\
\mathbf{x}^T_{i2} =(1,1), \qquad \forall i \\
z_{it} = 1 , \qquad \forall i, t \\
$$


$$
g(\mu_{it}) = \mathbf{x}^T_{it}\boldsymbol{\beta} + \mathbf{z}_{it}^T \mathbf{u}_i
\\
\begin{bmatrix} 
g(\mu_{i1}) \\ 
g(\mu_{i2}) \\ 
\end{bmatrix} = 
\begin{bmatrix} 
\mathbf{x}_{i1}^\top \\ 
\mathbf{x}_{i2}^\top \\ 
\end{bmatrix} \boldsymbol{\beta} + 
\begin{bmatrix} 
\mathbf{z}_{i1}^\top \\ 
\mathbf{z}_{i2}^\top \\ 
\end{bmatrix} \mathbf{u}_i
\\
\begin{bmatrix} 
g(\mu_{i1}) \\ 
g(\mu_{i2}) \\ 
\end{bmatrix} = 
\begin{bmatrix} 
1 & 0 \\ 
1 & 1 \\ 
\end{bmatrix} 
\begin{bmatrix} 
\alpha \\
\beta
\end{bmatrix} 
+ 
\begin{bmatrix} 
1 \\ 
1 \\ 
\end{bmatrix} \mathbf{u}_i
\\
= 
\begin{bmatrix} 
1 & 0 \\ 
1 & 1 \\ 
\end{bmatrix} 
\begin{bmatrix} 
\alpha \\
\beta
\end{bmatrix} 
+ 
\begin{bmatrix} 
u_i \\ 
u_i \\ 
\end{bmatrix} \\
= \begin{bmatrix} 
\alpha + u_i \\ 
\alpha + \beta + u_i \\ 
\end{bmatrix} 
$$


## Rasch Model and Item Response Models

(matched pair i induces the cluster, t can be thought of as the treatment)

- An extension of the logistic-match pairs model that allows $T>2$ observations in each cluster.
- The random intercept model has the form

$$
logit [P(\mathbf{Y}_{it}=1 \mid u_i)] = \alpha + \beta_t + u_i
$$




```{r}
summary(model_glm)
```

```{r}
summary(model_glmm )
```


```{r}
summary(model_glm)
```


```{r}
summary(model_glmm)
```


```{r}
model_glm  <- glm(outcome ~ drug + time + drug*time, family = binomial, data = depression_data)
model_glmm <- glmer(outcome ~ drug + time + drug*time + (1 | case), family = binomial, data = depression_data)
```

glmer(response ~ treat * time + (1 | id), family = binomial)



```{r}
model_glm$deviance - deviance(model_glmm)
```

```{r}
model_glm$df.residual - df.residual(model_glmm)
# 3.84
```




```{r}
Contraception %>% ggplot(aes(x = age, color = use)) +
  geom_boxplot() +
  facet_grid(fake_district~.) + theme_bw()
```






```{r}
# grid_long %>% filter(age == 0.0) %>% arrange(desc(prob_glmm))
```





```{r}
# # library(lme4)
# # 1. Fit the model
# m1 <- glmer(use ~ age +  (1 | fake_district), family = binomial, data = Contraception)
# # 2. Extract District Variance
# dist_var <- as.data.frame(VarCorr(m1))$vcov[1]
# 
# # 3. Apply formula (pi^2 / 3 ≈ 3.289)
# icc_value <- dist_var / (dist_var + (pi^2 / 3))
# print(icc_value)
```




```{r}
# summary(model_glm)
```


$$
g(\boldsymbol{\mu}_i) = \mathbf{X}_i \boldsymbol{\beta} + \mathbf{Z}_i \mathbf{u}_i
\\
\begin{bmatrix} 
g(\mu_{i1}) \\ 
g(\mu_{i2}) \\ 
\vdots \\ 
g(\mu_{iT_i}) 
\end{bmatrix} = 
\begin{bmatrix} 
\mathbf{x}_{i1}^\top \\ 
\mathbf{x}_{i2}^\top \\ 
\vdots \\ 
\mathbf{x}_{iT_i}^\top 
\end{bmatrix} \boldsymbol{\beta} + 
\begin{bmatrix} 
\mathbf{z}_{i1}^\top \\ 
\mathbf{z}_{i2}^\top \\ 
\vdots \\ 
\mathbf{z}_{iT_i}^\top 
\end{bmatrix} \mathbf{u}_i
$$









## III. Statistical Core (The "Master's" Details) – 8 Minutes

* The Estimation Challenge (The Integral): This is the biggest difference from GLMs. Explain that the marginal likelihood $L(\beta, \theta) = \int f(y|u)f(u)du$ has no closed-form solution.
* Approximation Methods: Briefly name-drop how software solves this: Laplace Approximation or Gauss-Hermite Quadrature.
* Shrinkage (Empirical Bayes): Explain that $u$ are not "estimated" like $\beta$, but "predicted" via Empirical Bayes. This causes Shrinkage—pulling unstable group estimates toward the global mean.
* Random Intercepts vs. Slopes: Use a visual to show how $Z$ changes to allow slopes to vary.

## IV. Implementation & Inference – 5 Minutes

* Software: glmer (lme4) or glmmTMB in R.
* Interpretation Paradox: Mention that $\beta$ in a GLMM is a subject-specific effect. (e.g., "For a specific person, a 1-unit increase in X increases their log-odds by $\beta$"). Contrast this briefly with GEE (Population-Averaged).
* Inference Issues: Mention the Boundary Problem: testing $H_0: \sigma^2 = 0$ is tricky because the null value is at the edge of the possible range (you can't have negative variance), making standard Likelihood Ratio Tests slightly conservative.

## V. Summary & Q&A – 5 Minutes

* Takeaway: GLMM = GLM + Random Effects + Integration.
* Pro-Tip: If you show a table of results, make sure to point out the Variance Components ($\sigma^2_u$) and the Residual Variance.


## Why these additions matter for your class:

   1. The $Z$ Matrix: In an MS Stats class, "The Equation" isn't complete without acknowledging the design matrix for the random effects.
   2. The Integral: This is the theoretical "heart" of why GLMMs are harder than GLMs. Acknowledging that we have to integrate out $u$ shows you understand the underlying calculus.

## 🕒 I. The Problem (The Data) – 5 mins

* Show the Data: Use a classic example (e.g., students in schools, or patients in clinics).
* The "Illegal" Step: Show a quick plot of the data. Point out that points within a group are closer to each other than to the rest.
* The GLM Failure: Explain that GLMs assume independence. If we ignore the groups, we commit Pseudoreplication.
* ⚡ Key Point: Tell them the GLM will give us "overconfident" (too small) p-values because it thinks we have more independent information than we actually do.

## 📐 II. The Equation (The Anatomy) – 8 mins

* Transition to Notation: Move from the data plot to the math.
* The Formula: $g(E[Y|u]) = X\beta + Zu$
* $X\beta$ (Fixed): The "Average" effect we care about.
   * $Zu$ (Random): The "Group-specific" noise.
* The Distribution: Mention $u \sim N(0, G)$. We aren't estimating every group's specific value; we are estimating the variance between them.

## 🧠 III. The "Master's" Hurdle (Integration) – 5 mins

* The No-Go Zone: Unlike a GLM, there is no closed-form solution.
* The Integral: Show the integral: $L = \int f(y|u)f(u)du$.
* The Explanation: To find the best $\beta$, we have to "average out" (integrate) all the possible values of the random effects $u$.
* The Tool: Mention we use Laplace Approximation to turn that scary integral into something a computer can optimize.

## 💻 IV. R Example & The "Fixed vs. Random" Debate – 10 mins

* The Comparison: Show a table comparing a glm() vs. a glmer().
* Point out how the Standard Errors get larger (and more honest) in the GLMM.
* The "District" Question: Ask the class: "Why not just put 'District' as a fixed factor?"
* Loss of Degrees of Freedom: If you have 50 districts, you waste 49 degrees of freedom.
   * Inference: Fixed effects only tell us about those 50 districts. Random effects let us generalize to all districts in the country.
   * Shrinkage: Mention that GLMMs "pull" weird outlier groups toward the global average (Empirical Bayes), making them more robust.

## 🏁 V. Wrap-up – 2 mins

* Summary: GLMM = The flexibility of GLM + The grouping power of Mixed Models.
* Final Tip: Check your variance components. If $\sigma^2$ is near zero, you might not even need a GLMM!

------------------------------
## Pro-Tip for your Live Demo:
When you show the R output, the class will likely ask about the "Correlation of Fixed Effects" table at the bottom. You can tell them that for a 30-minute intro, they can mostly ignore that—focus instead on the "Random Effects" block where it shows the Std.Dev. of the Intercept.
Would you like a sample R code snippet with a built-in dataset (like cbpp or grouseticks) that you can copy-paste for your demo?




