---
title: "GLMM"
output:
  ioslides_presentation:
    incremental: true
    widescreen: true
    smaller: true
    transition: faster
    toc: true
---

<style>
/* This targets the slide transition speed (default 'faster' is ~0.2s) */
slides > slide {
  -webkit-transition: all 0.1s ease-in-out;
  -moz-transition: all 0.1s ease-in-out;
  -o-transition: all 0.1s ease-in-out;
  transition: all 0.1s ease-in-out;
}
.scrollable-slide {
  overflow-y: auto !important;
  height: 500px;
}
.hidden-slide {
  display: none;
}
</style>


```{r setup, include=FALSE} 
knitr::opts_chunk$set(
  echo = FALSE, 
  comment = NA, 
  message = FALSE, 
  warning = FALSE,
  fig.align = 'center',
  fig.height = 4,
  fig.width = 6)
options(scipen = 1, digits = 5)
```

```{r}
depression_data <- read.table("Depression.dat", header = TRUE)
```


```{r}
library(mlmRev)
data(Contraception)
library(kableExtra)
library(dplyr)
library(tidyr)
library(MASS)
library(pROC)
library(caret)
library(knitr)
```

## Outline

- GLM Limitations  
- Random Effects Model
- Example

# GLM Limitations

## GLM

- Assumes independent observations

## Repeated measures {.build}

- The data shows multiple observations per individual (Case ID). Each individual has data on severity, drug, and outcome measurement in three different time points. (decription on page 456)

```{r}
library(DT)

subset_data <- depression_data %>% 
  filter(case %in% c(21, 22, 23))

datatable(subset_data, options = list(pageLength = 9,
                                      searching = FALSE, 
                                      lengthChange = FALSE, 
                                      dom = 't'))
```

- GLM would assume each row is independent from others, as if we have 9 independent samples.
- Observations within clusters are usually positively correlated, tending to be more alike than observations from different clusters

## Repeated measures {.build}

Reshaping highlights that we only have 3 independent samples, each with multiple observations.


```{r}
subset_data1 <- depression_data %>% filter(case %in% c(21, 22, 23)) %>% pivot_wider(names_from = 'time', values_from = 'outcome')

datatable(subset_data1, options = list(pageLength = 9,
                                      searching = FALSE, 
                                      lengthChange = FALSE, 
                                      dom = 't'))
```

## Natural groupings in data {.build}

- This shows yes-no responses on use of contraceptives among women of different ages from three out of 60 districts in Bangladesh.

```{r, fig.width=6, fig.height=3}
cont_filtered1 <- Contraception %>% filter(district %in% c(1, 34, 30))#c(34, 21, 61))#c(18, 30, 46)) 
cont_filtered2 <- cont_filtered1
cont_filtered2$district <- "combined"
rbind(cont_filtered1, cont_filtered2)%>% 
  ggplot(aes(x = district, y = age, color = use))+
  facet_grid(.~district, scales = "free_x") +
  geom_boxplot()  +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed")+ theme_bw() +
    theme(
    axis.text.x = element_blank(),  # Removes the labels (e.g., "18", "30")
    axis.ticks.x = element_blank(), # Removes the small tick marks
    axis.title.x = element_blank()  # Optional: Removes the "district" title
  )
  # group_by(district, use) %>% count()
```
- Ordinary analyses, such as GLM, that ignore the correlation and treat within-cluster observations the same as between-cluster observations produce invalid standard errors, tending to be too small for between-cluster effect estimates and too large for within-cluster effects.


## Natural groupings in data {.build}

```{r}
num_district <- Contraception %>% pull(district) %>% unique() %>% length()
```

- Why not use districts as one of the explanatory variables?
  - This dataset actually includes `r num_district` districts. Some of which have very few observations.
  
```{r, fig.width = 8, fig.height = 4}
Contraception %>% group_by(district) %>% count() %>% arrange(n) %>%
  ggplot(aes(x = reorder(district, desc(n)), y = n)) +
  geom_col() + theme_bw()  +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  #geom_hline(aes(yintercept = min(n)), color = "red", linetype = "dashed") +
  labs(title = 'Number of observations per district')
```

```{r}
dist <- Contraception %>% 
  group_by(district) %>% 
  count() %>% 
  arrange(desc(n)) %>% 
  head(10) %>% 
  pull(district)
```

## GLM

- Assumes independent observations
- Observations within clusters are usually positively correlated, tending to be more alike than observations from different clusters.  
- Ordinary analyses that ignore the correlation and treat within-cluster observations the same as between-cluster observations produce invalid standard errors, tending to be too small for between-cluster effect estimates and too large for within-cluster effects.
- Solution: add cluster-level terms to the model that take the same value for each observation in a cluster. They are unobserved and when treated as varying randomly among clusters, they are called random effects.

# Random Effects Model

```{r}
# Contraception %>% #filter(district %in% c(18, 30, 46)) %>% 
#   ggplot(aes(x = factor(fake_district), y = age, color = use))+
#   facet_grid(.~fake_district, scales = "free_x") +
#   geom_boxplot()  +
#   geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed")+ theme_bw()
```



## Definitions {.build}

- **Fixed effects** are parameters that describe a factor's effects in an ordinary GLM. They apply to all categories of interest, such as gender, treatment, or age groupings.
- **Random effects** usually apply to a sample. For a study that makes repeated observations on a sample of subjects, the model treats observations from a given subject as a cluster, and it has a random effect for each of the subject.
- GLMs extend ordinary regression by allowing nonnormal responses and a link function of the mean. The GLMM is a further extension that permits random effects as well as fixed effects in a linear predictor.

## GLMM {.build}

- you have $I$ clusters: $i = 1, \dots, I$ with $T_i$ observations in the $i$th cluster
- $y_{it}$: observation $t$ in cluster $i$
- the number of observations may vary by cluster
- $\mathbf{x}_{it}$: column vector of values of explanatory variables for $y_{it}$
- $\mathbf{u}_i$: vector of random effects for cluster $i$. This is often univariate. This is assumed to have a multivariate normal distribution $\mathcal{N}(\mathbf{0}, \boldsymbol{\Sigma})$. The covariance matrix $\boldsymbol{\Sigma}$ depends on unknown variance components and possible also correlation parameters.
- Conditional on $\mathbf{u}_i$, a GLMM resembles an ordinary GLM. Conditional on $\mathbf{u}_i$, standard model fitting treats $\{y_{it}\}$ as independent over $t$ and $i$.
- Let $\mu_{it} = E(Y_{it} \mid \mathbf{u}_i)$. The linear predictor for a GLMM for link function $g(\cdot)$ and fixed effect model parameters $\boldsymbol{\beta}$ has the form

$$
g(\mu_{it}) = \mathbf{x}^T_{it}\boldsymbol{\beta} + \mathbf{z}_{it}^T \mathbf{u}_i
$$



## GLMM - Random Intercept {.build .scrollable-slide}

Let $T = 2$, and $\boldsymbol{\beta} = \begin{bmatrix} \alpha \\ \beta\end{bmatrix}$.

For a single observation $(i, t)$, the formula is

$$
g(\mu_{it}) = \mathbf{x}^T_{it}\boldsymbol{\beta} + \mathbf{z}_{it}^T \mathbf{u}_i
$$

If $\mathbf{u}_i$ is univariate, we can either have a random slope or a random intercept. Suppose we have a random intercept. Note that $u_i \sim\mathcal{N}(0, \sigma^2)$

$$
\begin{aligned}
\begin{bmatrix} 
g(\mu_{11}) \\ 
g(\mu_{12}) \\ 
g(\mu_{21}) \\ 
g(\mu_{22}) \\ 
\vdots \\
g(\mu_{I1}) \\ 
g(\mu_{I2}) \\ 
\end{bmatrix} 
&= 
\begin{bmatrix} 
\mathbf{x}_{11}^\top \\ 
\mathbf{x}_{12}^\top \\ 
\mathbf{x}_{21}^\top \\ 
\mathbf{x}_{22}^\top \\ 
\vdots \\
\mathbf{x}_{I1}^\top \\ 
\mathbf{x}_{I2}^\top \\ 
\end{bmatrix}
\begin{bmatrix} \alpha \\ \beta\end{bmatrix}
+ 
\begin{bmatrix} 
\mathbf{z}_{11}^\top \\ 
\mathbf{z}_{12}^\top \\ 
\mathbf{z}_{21}^\top \\ 
\mathbf{z}_{22}^\top \\ 
\vdots \\
\mathbf{z}_{I1}^\top \\ 
\mathbf{z}_{I2}^\top \\ 
\end{bmatrix}
\begin{bmatrix} u_1 \\ u_2 \\ \vdots \\ u_I \end{bmatrix} \\
& = 
\begin{bmatrix} 
1 & x_{11} \\ 
1 & x_{12} \\ 
1 & x_{21} \\ 
1 & x_{22} \\ 
\vdots \\
1 & x_{I1} \\ 
1 & x_{I2} \\ 
\end{bmatrix}
\begin{bmatrix} \alpha \\ \beta\end{bmatrix} 
+ 
\begin{bmatrix} 
1 & 0 & \cdots & 0 \\ 
1 & 0 & \cdots & 0 \\ 
0 & 1 & \cdots & 0 \\ 
0 & 1 & \cdots & 0 \\ 
\vdots & \vdots & \ddots & \vdots \\
0 & 0 & \cdots & 1 \\ 
0 & 0 & \cdots & 1 \\ 
\end{bmatrix}
\begin{bmatrix} u_1 \\ u_2 \\ \vdots \\ u_I \end{bmatrix} \\
g[E(\mathbf{Y} \mid \mathbf{u})] = g(\boldsymbol{\mu}) 
&= \mathbf{X}_{(2I)\times 2} \boldsymbol\beta_{2 \times 1} 
+ \mathbf{Z}_{(2I)\times I} \mathbf{u}_{I \times 1}, 
\qquad \text{matrix notation}
\\
\begin{bmatrix} 
g(\mu_{11}) \\ 
g(\mu_{12}) \\ 
g(\mu_{21}) \\ 
g(\mu_{22}) \\ 
\vdots \\
g(\mu_{I1}) \\ 
g(\mu_{I2}) \\ 
\end{bmatrix}
&=
\begin{bmatrix} 
(\alpha + u_1) + x_{11}\beta \\ 
(\alpha + u_1) + x_{12}\beta \\ 
(\alpha + u_2) + x_{21}\beta \\ 
(\alpha + u_2) + x_{22}\beta \\ 
\vdots \\
(\alpha + u_I) + x_{I1}\beta \\ 
(\alpha + u_I) + x_{I2}\beta \\ 
\end{bmatrix}
\end{aligned}
$$

- Link(Average outcome) = General trend (Fixed) + Cluster deviation (Random)
- $\alpha$ and $\beta$ are the fixed effects while $u_1$ and $u_2$ are the random effects realization.
- If $Y_{it} \mid u_i \sim \text{Bernoulli}(\mu_{it})$, $g = \text{logit}$, and $x_{it}$ indicates whether treatment is administered or not, we have a logistic GLMM with random intercept for binary matched pairs.
  - an extension allows $T > 2$ observations in each cluster. The random intercept model has form below and early applications were in psychometrics (item response models).
  
  $$
  \text{logit}[P(Y_{it}=1 \mid u_i)] = \alpha + \beta_t + u_i
  $$
    - The item-response model with random intercept is a special case of an important class of random effects model for binary data called logistic-normal models for binary data with the form
    
    $$
    \text{logit}[P(Y_{it}=1 \mid u_i)] = \mathbf{x}^T_{it} \boldsymbol{\beta}+ u_i
    $$


## Heterogeneity in Logistic-Normal models {.build .scrollable-slide}


Take 2 observations from different clusters. Let them be $(x_{it}, y_{it})$ and $(x_{hs}, y_{hs})$. We look at their difference in log-odds.

$$
\text{logit}[P(Y_{it}=1 \mid u_i)] = \mathbf{x}^T_{it} \boldsymbol{\beta}+ u_i
$$

$$
\text{logit}[P(Y_{hs}=1 \mid u_h)] = \mathbf{x}^T_{hs} \boldsymbol{\beta}+ u_h
$$


$$
\begin{aligned}
\text{logit}[P(Y_{it}=1 \mid u_i)] - \text{logit}[P(Y_{hs}=1 \mid u_h)] &=
\mathbf{x}^T_{it} \boldsymbol{\beta} + u_i - \mathbf{x}^T_{hs} \boldsymbol{\beta} - u_h \\
&= (\mathbf{x}^T_{it}  - \mathbf{x}^T_{hs}) \boldsymbol{\beta} + (u_i - u_h) \\
&= (\mathbf{x}_{it}  - \mathbf{x}_{hs})^T \boldsymbol{\beta} + (u_i - u_h)
\end{aligned}s
$$


- Fixed part: $(\mathbf{x}_{it}  - \mathbf{x}_{hs})^T \boldsymbol{\beta}$
- Random part: $(u_i - u_h) \sim \mathcal{N}(0, 2 \sigma^2)$
- $u_i - u_h$ captures additional variation in log-odds due to cluster-level heterogeneity.

If $\sigma = 0$, there is no variation across clusters - the logistic-normal model simplifies to the ordinary logistic regression model, treating all observations as independent.

$$
\begin{aligned}
\text{logit}[P(Y_{it}=1 \mid u_i)] - \text{logit}[P(Y_{hs}=1 \mid u_h)]
&= (\mathbf{x}_{it}  - \mathbf{x}_{hs})^T \boldsymbol{\beta} + (u_i - u_h) \\
&= (\mathbf{x}_{it} - \mathbf{x}_{hs})^T \boldsymbol{\beta}
\end{aligned}
$$

If $\sigma > 0$, $(\mathbf{x}_{it}  - \mathbf{x}_{hs})^T \boldsymbol{\beta} + (u_i - u_h)$ only reduces to $(\mathbf{x}_{it} - \mathbf{x}_{hs})^T \boldsymbol{\beta}$ if $u_i = u_h$ --- when observations come from the same cluster.

## Form: GLM vs GLMM {.big-text .build}

Suppose we use the logit link:

GLM

$$
\text{logit}[P(Y_{it}=1)] = \mathbf{x}^T_{it} \boldsymbol{\beta}
$$

GLMM; $u_i \sim \mathcal{N}(0,\sigma^2)$

$$
\text{logit}[P(Y_{it}=1 \mid u_i)] = \mathbf{x}^T_{it} \boldsymbol{\beta}+ u_i
$$


## Random effects: unmeasured predictors {.build}

The random effects sometimes represent heterogeneity caused by omitting certain explanatory variables. Suppose we only have $p-1$ predictors,

$$
g[E(\mathbf{y})] = g(\boldsymbol{\mu}) = \mathbf{X}_{n \times (p-1)} \boldsymbol{\beta}_{(p-1)\times 1}
$$

But the model could have been better explained in the presence of the $p$th variable which is unmeasured.

The random effects model relate to methods of dealing with unmeasured predictors. The random effects part of the linear predictor reflects terms that would be in the fixed effects part if those explanatory variables had been included.

$$
g[E(\mathbf{y})] = \mathbf{X}_{n \times (p-1)} \boldsymbol{\beta}_{(p-1)\times 1} + \mathbf{Z}_{n \times q}\mathbf{u}_{q \times 1}
$$

## Random effects: random measurement error in the explanatory variables {.build}

Random effects also sometimes represent random measurement error in the explanatory variables. 

If we replace a particular $x_{it}$ by $x^*_{it} + \epsilon_i$, with $x^*_{it}$ the true values and $\epsilon_i$ the measurement error, then $\epsilon_i\beta$ can be absorbed in the random effects term.


$$
\begin{aligned}
g(\mu_{it}) &= \mathbf{x}^T_{it}\boldsymbol{\beta} + \mathbf{z}_{it}^T \mathbf{u}_i
\\
&= (\mathbf{x}^{T*}_{it} + \epsilon_i)\boldsymbol{\beta} + \mathbf{z}_{it}^T \mathbf{u}_i
\\
&= \mathbf{x}^{T*}_{it}\boldsymbol{\beta} + \epsilon_i\boldsymbol{\beta} + \mathbf{z}_{it}^T \mathbf{u}_i
\end{aligned}
$$


## Random effects: explaining overdispersion {.build}

Overdispersion occurs when the model underestimates the variance in the outcome; it has the ff implications  

- standard errors that are too small  
- p-values biased towards significance  
- CI too narrow  

Potential sources of overdispersion: 

- observations are not independent  
- external factors influencing the outcome, not included in the model  

Random effects provide a mechanism for explaining overdispersion in basic models not having those effects.


# Example

## Dataset

```{r}
library(mlmRev)
data(Contraception)

datatable(Contraception, options = list(pageLength = 9,
                                      #searching = FALSE, 
                                      lengthChange = FALSE, 
                                      dom = 't'))
```

## Fit a GLM {.build .scrollable-slide}

We ignore the district variable.

```{r echo = TRUE}
model_glm  <- glm(use ~ age + urban + livch, 
                  family = binomial, 
                  data = Contraception)
```

```{r}
summary(model_glm)
```


## Fit a GLMM {.build .scrollable-slide}

We use random intercept model.

```{r echo = TRUE}
model_glmm0 <- glmer(use ~ age + urban + livch + (1 | district), 
                    family = binomial, 
                    data = Contraception)
```

```{r}
summary(model_glmm0)
```


```{r echo = TRUE}
ranef(model_glmm0)
```

## Visualize the characteristic curve

```{r fig.height = 5, fig.width = 7}
long_age_range <- seq(-250, 250, length.out = 1201)
districts <- c(1, 30, 34)#c(18, 30, 46)#unique(Contraception$district)
fake_districts <- c(1, 2, 3)
urban <- 'Y'#unique(Contraception$fake_district)
livch <- 1
grid_long <- expand.grid(age = long_age_range, 
                         district = districts,
                         livch = livch,
                         urban = urban)
grid_long <- grid_long %>% mutate(district = factor(district),
                                  livch = factor(livch))

grid_long$prob_glmm <- predict(model_glmm0, newdata = grid_long, type = "response", re.form = NULL)
grid_long$prob_glm  <- predict(model_glm,  newdata = grid_long, type = "response")

ggplot(grid_long, aes(x = age)) +

  
  geom_line(aes(y = prob_glmm, group = district, linetype = district),
            color = "blue", alpha = 0.5) +

  geom_line(aes(y = prob_glm,color = "GLM"),
            color = "red", size = 0.5) +
  scale_linetype_manual(values = c("longdash","solid", "dotted")) + 

  geom_hline(yintercept = 0.5, linetype = "dotted", alpha = 0.5) +
  geom_vline(xintercept = 0, linetype = "dotted", alpha = 0.5) +
  theme_minimal() +
  labs(title = "GLM vs. GLMM",
       subtitle = "Red: GLM | Blue: GLMM Districts",
       x = "Age (Centered/Scaled)",
       y = "Predicted Probability") + theme_bw()
```


## What happens if {.build .scrollable-slide}

... we fit a GLMM to a dataset but the clusters do not vary? 

It gets reduced to a GLM. 

```{r}
set.seed(658) # still diff
set.seed(7)
Contraception <- Contraception[sample(nrow(Contraception)), ]
Contraception$fake_district <- c(rep(1, 645), rep(2, 645), rep(3, 644))
```

### `age` per district

```{r}
Contraception %>% 
  ggplot(aes(x = fake_district, y = age, color = use))+
  facet_grid(.~fake_district, scales = "free_x") +
  geom_boxplot()  +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed")+ theme_bw() +
    theme(
    axis.text.x = element_blank(),  # Removes the labels (e.g., "18", "30")
    axis.ticks.x = element_blank(), # Removes the small tick marks
    axis.title.x = element_blank()  # Optional: Removes the "district" title
  )
```

### `livch` per district

```{r}
Contraception %>% 
  group_by(fake_district, livch) %>% 
  summarise(prop_using = mean(as.character(use) == "Y")) %>%
  ggplot(aes(x = livch, y = prop_using)) +
  geom_col() +
  facet_grid(~fake_district) +
  theme_bw()
```

### `urban` per district

```{r}
Contraception %>% 
  group_by(fake_district, urban) %>% 
  summarise(prop_using = mean(as.character(use) == "Y")) %>%
  ggplot(aes(x = urban, y = prop_using)) +
  geom_col() +
  facet_grid(~fake_district) +
  theme_bw()
```

### Fit a GLMM

```{r echo = TRUE}
model_glmm1 <- glmer(use ~ age + urban + livch + (1 | fake_district), 
                    family = binomial, 
                    data = Contraception)
```


```{r}
summary(model_glmm1)
```

```{r echo = TRUE}
ranef(model_glmm1)
```


### Visualize the characteristic curve

Note that the GLM characteristic curve and the three curves corresponding to the 3 fake districts overlap.

When there is no between-cluster variability ($\sigma^2 =0$), the GLMM reduces to a GLM.

```{r fig.height = 5, fig.width = 7}
grid_long <- expand.grid(age = long_age_range, 
                         fake_district = fake_districts,
                         livch = livch,
                         urban = urban)
grid_long <- grid_long %>% mutate(district = factor(fake_district),
                                  livch = factor(livch))

grid_long$prob_glmm <- predict(model_glmm1, newdata = grid_long, type = "response", re.form = NULL)
grid_long$prob_glm  <- predict(model_glm,  newdata = grid_long, type = "response")

ggplot(grid_long, aes(x = age)) +
  geom_line(aes(y = prob_glmm, group = fake_district, linetype = district),
            color = "blue", alpha = 0.25) +

  geom_line(aes(y = prob_glm,color = "GLM"),
            color = "red", alpha = 1, linetype = "dotted") +
  scale_linetype_manual(values = c("longdash","solid", "dotted")) + 

  geom_hline(yintercept = 0.5, linetype = "dotted", alpha = 0.5) +
  geom_vline(xintercept = 0, linetype = "dotted", alpha = 0.5) +
  theme_minimal() +
  labs(title = "GLM vs. GLMM",
       subtitle = "Red: GLM | Blue: GLMM Fake Districts",
       x = "Age (Centered/Scaled)",
       y = "Predicted Probability") + theme_bw()
```






# Appendix

## GLMM - Random Slope {.build .scrollable-slide}

Let $I = 2$, $T = 2$, and $\boldsymbol{\beta} = \begin{bmatrix} \alpha \\ \beta\end{bmatrix}$.

For a single observation $(i, t)$, the formula is

$$
g(\mu_{it}) = \mathbf{x}^T_{it}\boldsymbol{\beta} + \mathbf{z}_{it}^T u_i
$$

If $u_i$ is a **random slope** for the covariate $x_{it}$, then:

$$
\begin{aligned}
\begin{bmatrix} 
g(\mu_{11}) \\ 
g(\mu_{12}) \\ 
g(\mu_{21}) \\ 
g(\mu_{22}) \\ 
\end{bmatrix} &= 
\begin{bmatrix} 
1 & x_{11} \\ 
1 & x_{12} \\ 
1 & x_{21} \\ 
1 & x_{22} \\ 
\end{bmatrix}
\begin{bmatrix} \alpha \\ \beta\end{bmatrix}
+ 
\begin{bmatrix} 
x_{11} & 0 \\ 
x_{12} & 0 \\ 
0 & x_{21} \\ 
0 & x_{22} \\ 
\end{bmatrix}
\begin{bmatrix} u_1 \\ u_2\end{bmatrix} \\
\\
\begin{bmatrix} 
g(\mu_{11}) \\ 
g(\mu_{12}) \\ 
g(\mu_{21}) \\ 
g(\mu_{22}) \\ 
\end{bmatrix}
&=
\begin{bmatrix} 
\alpha + x_{11}(\beta + u_1) \\ 
\alpha + x_{12}(\beta + u_1) \\ 
\alpha + x_{21}(\beta + u_2) \\ 
\alpha + x_{22}(\beta + u_2) \\ 
\end{bmatrix}
\end{aligned}
$$

- The term $(\beta + u_i)$ represents the cluster-specific slope.
- $\alpha$ and $\beta$ are the fixed effects while $u_1$ and $u_2$ are the random slope deviations.

## GLMM - Random Intercept & Slope {.build .scrollable-slide}

Let $I = 2$, $T = 2$, and $\boldsymbol{\beta} = \begin{bmatrix} \alpha \\ \beta\end{bmatrix}$. 

Now, $\mathbf{u}_i = \begin{bmatrix} u_{0i} \\ u_{1i} \end{bmatrix}$ is a **vector** (bivariate) containing a random intercept and a random slope.

$$
\begin{aligned}
\begin{bmatrix} 
g(\mu_{11}) \\ 
g(\mu_{12}) \\ 
g(\mu_{21}) \\ 
g(\mu_{22}) \\ 
\end{bmatrix} &= 
\begin{bmatrix} 
1 & x_{11} \\ 
1 & x_{12} \\ 
1 & x_{21} \\ 
1 & x_{22} \\ 
\end{bmatrix}
\begin{bmatrix} \alpha \\ \beta\end{bmatrix}
+ 
\begin{bmatrix} 
1 & x_{11} & 0 & 0 \\ 
1 & x_{12} & 0 & 0 \\ 
0 & 0 & 1 & x_{21} \\ 
0 & 0 & 1 & x_{22} \\ 
\end{bmatrix}
\begin{bmatrix} u_{01} \\ u_{11} \\ u_{02} \\ u_{12} \end{bmatrix} \\
\\
\begin{bmatrix} 
g(\mu_{11}) \\ 
g(\mu_{12}) \\ 
g(\mu_{21}) \\ 
g(\mu_{22}) \\ 
\end{bmatrix}
&=
\begin{bmatrix} 
(\alpha + u_{01}) + x_{11}(\beta + u_{11}) \\ 
(\alpha + u_{01}) + x_{12}(\beta + u_{11}) \\ 
(\alpha + u_{02}) + x_{21}(\beta + u_{12}) \\ 
(\alpha + u_{02}) + x_{22}(\beta + u_{12}) \\ 
\end{bmatrix}
\end{aligned}
$$

- **Fixed Part:** $\alpha$ (intercept) and $\beta$ (slope).
- **Random Part:** $u_{0i}$ (intercept deviation) and $u_{1i}$ (slope deviation) for each cluster $i$.
- $\mathbf{Z}$ is now a **block-diagonal matrix** where each block is the $\mathbf{z}_{it}$ values for that cluster.