## Data source

```{r}
raw <- read.csv("data/Replication file/Data/G-SWA.csv") %>% 
  filter(country == 'Australia 2') %>% 
  dplyr::select(c(regionname,
                  gender,
                  age, #agegroups,
                  education,
                  industry_job,
                  married,
                  with_kids,
                  daysemployer_work_home,
                  daysemployee_work_home,
                  value_WFH_rawpercent25,
                  commuting_time, # * commute_time_hs
                  WFHperceptions,
                  # ever_WFH,
                  return_office
                  )) %>%
  # dplyr::select(-c(daysemployer_work_home)) %>%
  mutate(decision = ifelse(return_office == 'Comply and return to worksite', 0, 1)) %>%
 # mutate(employee_favored_agreement = daysemployee_work_home - daysemployer_work_home) %>%
  dplyr::select(-c(daysemployer_work_home,return_office))

# raw[apply(raw, 1, function(x) any(is.na(x) | x == "")), ]
```




```{r}
n <- dim(raw)[1]
```

This study uses the Global Survey of Working Arrangements (G-SWA, 2022) dataset. The G-SWA is a survey of individual workers across more than two dozen countries done occasionally. It collects data on outcomes, plans, desires and perceptions related to WfH along with the respondents' demographics, earnings, commute times, and more (@datasource-2022). We focus on the working class of Australia who primarily WfH and use the wave 2 phase of data collection. The total sample size is `r n`. The variables used in the analysis are defined as follows:

\begingroup
\setlength{\parindent}{0pt}

### Target variable

**`decision`** is a binary variable. This is derived from `return_office` which pertains to the response if employer announces that employees must return to the worksite 5+ days per week. `return_office` could be one of the three options: `Look for a job to WFH 1-2 days`, `Comply and return to worksite`, `Quit job`. For this study, we collapse `Look for a job to WFH 1-2 days` and `Quit job` to obtain the `decision` variable which is 0 if response is `Comply and return to worksite` and 1 otherwise. (Survey question: "How would you respond if
your employer announced that all employees must return to the worksite 5+ days a week, starting on February 1, 2022?")

```{r}
get_target_contingency <- function(raw_data, 
                                   explanatory_variable, 
                                   target_variable){
  return(
    raw_data %>% 
      group_by(.data[[explanatory_variable]]) %>%
      dplyr::count(.data[[target_variable]]) %>% 
      pivot_wider(names_from = target_variable, values_from = n) %>%
      # Calculate row total first
      mutate(row_total = sum(`0`, `1`, na.rm = TRUE)) %>%
      # Convert counts to proportions
      mutate(across(c(`0`, `1`), ~ . / row_total)) %>%
      arrange(desc(row_total))
  )
}

make_latex_table <- function(df) {
  df %>%
    kable("latex", 
          booktabs = TRUE,
          escape = TRUE,
          linesep = "",
          row.names = FALSE,
          na.character = "") %>%
    kable_styling(latex_options = "HOLD_position", 
                  position = "center", 
                  full_width = FALSE)
}
```


```{r}
target_var_tab <- raw %>% 
  dplyr::count(decision) %>%
  mutate(prop = round(n/sum(n), 4))

target_var_tab %>% make_latex_table()
```


### Categorical Explanatory variables

**`daysemployee_work_home`** is a count variable from 0 to 5, indicating the respondent's desired number of WfH days per week. (Survey question: "After COVID, in 2022 and later, how often would you like to work from home?")

```{r}
# get_target_contingency(raw, 
#                        "ever_WFH", 
#                        "decision") %>% make_latex_table()
```


```{r}
get_target_contingency(raw, 
                       "daysemployee_work_home", 
                       "decision") %>% make_latex_table()
```

**`education`** describes the respondent's educational attainment.

```{r}
get_target_contingency(raw, 
                       "education", 
                       "decision") %>% make_latex_table()
```

**`gender`** can either drop `Other/Prefer not to say` category due to its low representation or replace it with `male`, which is the dominant category in terms of `gender` frequency.

```{r}
get_target_contingency(raw, 
                       "gender", 
                       "decision") %>% make_latex_table()
```

**`industry_job`** indicates the respondent's industry of current or most recent job. We will further collapse these categories due to low frequencies.

```{r}
get_target_contingency(raw, 
                       "industry_job", 
                       "decision") %>% make_latex_table()
```


**`married`** is binary variable that is 1 if respondent is married and 0 otherwise.

```{r}
get_target_contingency(raw, 
                       "married", 
                       "decision") %>% make_latex_table()
```

**`regionname`** represents the respondent's subnational region. To avoid potential complete separation due to the number of categories (8), we may collapse them based on proximity (BCARR ring) or CBD status. If category grouping isn't feasible, we will treat `regionname` as a random effect instead. For the last row where `regionname` is unspecified, we will use the mode (`New Sount Wales`).

```{r}
get_target_contingency(raw, 
                       "regionname", 
                       "decision") %>% make_latex_table()
```

**`value_WFH_rawpercent25`** can be considered ordered categorical and is described as the amenity value of the option to WfH 2-3 days per week. We will try to collapse this into fewer categories---potentially, if it's negative, zero, or positive. Alternatively, we can collapse all negatives and zero (collapsed into not willing to pay) and describe the percent of the respondent's salary he or she is willing to pay in exchange of 2 or 3 WfH days. (Survey question: "How much of a pay raise [cut] (as a percent of your current pay) would you value as much as the option to work from home 2 or 3 days a week?")

```{r}
get_target_contingency(raw, 
                       "value_WFH_rawpercent25", 
                       "decision") %>%
  arrange(value_WFH_rawpercent25) %>% 
  make_latex_table()
```



**`WFHperceptions`** is an ordinal variable that describes the how WfH is being perceived now vs before the pandemic. Possible responses are listed below. Similar to other variables, we can collapse negatives and zero due to lack of representative frequencies for some negatives. (Survey question: Since the COVID pandemic began, how have perceptions about WFH changed among people you know?)

- Improved [among almost all (95%), most (70%) or some (25%)]
- No change (0%)
- Worsened [among almost all (-95%), most (-70%) or some (-25%)]

```{r}
get_target_contingency(raw, 
                       "WFHperceptions", 
                       "decision") %>%
  arrange(WFHperceptions) %>% 
  make_latex_table()
```

**`with_kids`** is binary variable that is 1 if respondent lives with children under 14 of years in age and 0 otherwise. For the `NA` below, we assume a value of 0 since the respondent indicated 0 under `married` variable.

```{r}
get_target_contingency(raw, 
                       "with_kids", 
                       "decision")  %>% make_latex_table()
```


### Continuous Explanatory variables

**`age`** is specified in units of years. We can see from the boxplot that the median `age` for those who intend to stay and comply when given a mandate to work in-office 100% is higher than those who would rather find another job. 

```{r fig.height=2, fig.width = 3}
library(ggplot2)
# raw %>% 
#   group_by(age) %>%
#   dplyr::count(decision) #%>%
#   #mutate(prop = round(n/sum(n), 4))

raw %>%
  ggplot(aes(x = age,y=factor(decision))) +
  geom_boxplot(position = "dodge", alpha = 0.5) + 
  # facet_wrap(~decision) +
  theme_bw() +
  labs(title = "Decision vs Age",
       x = "age (years)", 
       y = "decision")
```


```{r}
raw %>% 
  group_by(decision) %>% 
  summarise(as_tibble(t(as.matrix(summary(age)))), .groups = "drop") %>% 
  make_latex_table()
```



**`commuting_time`** is the respondent's round trip commute time in minutes. The boxplot show that the median time of respondents who would quit or find another job is higher than those who would choose to stay. However, there are a number of outliers who would still choose to stay despite their extremely long round trip commute time. We will assume that there is no error in the data --- the maximum is 480 mins (8 hours). (Survey question: "How long do you usually spend commuting to and from work (in minutes). If you are not currently commuting to work, please answer based on your commute time in 2019 (before COVID)")

```{r fig.height=2, fig.width = 3}
target_var_tab <- raw %>% 
  group_by(commuting_time) %>%
  dplyr::count(decision) #%>%
  #mutate(prop = round(n/sum(n), 4))

# raw %>%
#   ggplot(aes(x = commuting_time, fill = factor(decision))) +
#   geom_histogram(position = "identity", alpha = 0.5) + theme_bw()

raw %>%
  ggplot(aes(x = commuting_time,y=factor(decision))) +
  geom_boxplot(position = "dodge", alpha = 0.5) + 
  # facet_wrap(~decision) +
  theme_bw() +
  labs(title = "Decision vs Commuting Time",
       x = "commuting_time (roundtrip in mins)", 
       y = "decision")
```

```{r}
raw %>% 
  group_by(decision) %>% 
  summarise(as_tibble(t(as.matrix(summary(commuting_time)))), .groups = "drop") %>% 
  make_latex_table()
```

\endgroup

## Model specification

We want to consider a starting binary logistic regression model with all enumerated variables as main effects and interaction on the following variables: `gender` and `education`, `gender` and `industry_job`, `with_kids` and `gender`. We then implement stepwise regression and a series of likelihood ratio tests to come up with the final model. 

In the event of complete separation, we may try to consider firth logistic regression.

We will also try to explore considering `regionname` as random effect if we cannot find a way to effectively regroup the categories. 