---
title: "Your Thesis Title Here"
author: "Shaine Rosewel Matala"
date: "`r Sys.Date()`"
abstract: |
  Working From Home (WFH) has become the trend for many industries especially after the pandemic. 
number_sections: true
fig_caption: yes
output:
  bookdown::pdf_document2:
    keep_tex: true
toc: false
indent: true
header-includes:
- \usepackage{amsmath}
- \numberwithin{equation}{section}
- \numberwithin{figure}{section}
- \numberwithin{table}{section}
- \usepackage{setspace}\onehalfspacing
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
- \usepackage{indentfirst}
- \usepackage{xparse}
- \usepackage{tcolorbox}
- \usepackage{algorithm}
- \usepackage{algpseudocode}
- \usepackage{float}
- \floatplacement{figure}{H}
- \usepackage{placeins}
- \renewcommand{\ttfamily}{\rmfamily}
- \usepackage{etoolbox}
- \usepackage{multirow}
- \usepackage{makecell}
- \usepackage{booktabs}
- \AtBeginEnvironment{algorithmic}{\setstretch{1.5}} 
- \renewcommand{\arraystretch}{0.75}
- \renewcommand{\multirowsetup}{\raggedright}
- \usepackage{anyfontsize}
- \newcommand{\thesisfont}{\fontsize{12pt}{12pt}\selectfont}
- \thesisfont
- \usepackage{xcolor}
# Combined hyperref into one clean call to avoid conflicts
- \usepackage[colorlinks=true,linkcolor=blue,urlcolor=black,citecolor=black]{hyperref}
- \usepackage{titlesec}
- \titleformat{\section}{\fontsize{20pt}{22pt}\bfseries}{\thesection}{1em}{}
- \titleformat{\subsection}{\fontsize{18pt}{20pt}\bfseries}{\thesubsection}{1em}{}
- \titleformat{\subsubsection}{\fontsize{16pt}{19pt}\bfseries}{\thesubsubsection}{1em}{}
- \usepackage{caption}
- \captionsetup[figure]{font={large,bf}}
- \captionsetup[table]{font={large,bf}}
- \newcommand{\algfont}{\fontsize{15pt}{18pt}\selectfont}
papersize: a4
bibliography: child/references/references.bib
nocite: | 
  @*
csl: child/references/apa-6th-edition.csl
---



```{r echo = FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  message = FALSE,
  warning = FALSE,
  knitr.kable.NA = ''
)
```


# Introduction {#sec:intro}


```{r, child = "child/01_introduction.Rmd", echo = FALSE}

```

# Related Literature


```{r, child = "child/02_rrl.Rmd", echo = FALSE}

```

# Methodology

```{r, child = "child/03_methodology.Rmd"}

```

# Result

```{r, child = "child/04_result.Rmd"}

```

# Conclusion

```{r, child = "child/05_conclusion.Rmd"}

```

\clearpage
\vspace*{1cm}
\phantomsection

# Bibliography {-}

\vspace*{0.8cm}

::: {#refs}
:::
  
\clearpage
\vspace*{1cm}

\phantomsection 
\appendix


# Appendix {-}

\vspace*{0.8cm}

\setcounter{section}{0}
\renewcommand{\thesection}{A}

\setcounter{subsection}{0}
\renewcommand{\thesubsection}{A.\arabic{subsection}}

\numberwithin{table}{subsection}
\numberwithin{figure}{subsection}

```{r, child = "child/06_appendix.Rmd"}
```