---
title: "Modeling Workers' Return-to-Office Decisions in Australia"
author: "Shaine Rosewel Matala"
date: "`r Sys.Date()`"
abstract: |
  This paper uses the Global Survey of Working Arrangements (G-SWA, 2022) to describe respondents from the working class of Australia based on the decision they would take if their employer announced that all employees must return to the worksite 5+ days a week by a certain date. The response signals a preference for either a hybrid/WFH or 100% in-office work setting. We aim to describe this decision based on their immediate environment's perception of remote work, commute time, gender, age, whether they have children, industry, and other factors. We consider a generalized linear model with an identity or logistic link. This may be beneficial for employers and the government in understanding workers' needs and potentially informing regulations.
number_sections: true
fig_caption: yes
output:
  bookdown::pdf_document2:
    keep_tex: true
toc: false
indent: true
colorlinks: true
linkcolor: blue
urlcolor: blue
citecolor: blue
link-citations: true
header-includes:
- \usepackage{amsmath}
- \numberwithin{equation}{section}
- \numberwithin{figure}{section}
- \numberwithin{table}{section}
- \usepackage{setspace}\onehalfspacing
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
- \usepackage{indentfirst}
- \usepackage{xparse}
- \usepackage{tcolorbox}
- \usepackage{algorithm}
- \usepackage{algpseudocode}
- \usepackage{float}
- \floatplacement{figure}{H}
- \usepackage{placeins}
- \renewcommand{\ttfamily}{\rmfamily}
- \usepackage{etoolbox}
- \usepackage{multirow}
- \usepackage{makecell}
- \usepackage{booktabs}
- \AtBeginEnvironment{algorithmic}{\setstretch{1.5}} 
- \renewcommand{\arraystretch}{0.75}
- \renewcommand{\multirowsetup}{\raggedright}
- \usepackage{anyfontsize}
- \newcommand{\thesisfont}{\fontsize{12pt}{12pt}\selectfont}
- \thesisfont
- \usepackage{xcolor}
# Combined hyperref into one clean call to avoid conflicts
- \usepackage{titlesec}
- \usepackage{caption}
- \captionsetup[figure]{font={large,bf}}
- \captionsetup[table]{font={large,bf}}
- \newcommand{\algfont}{\fontsize{15pt}{18pt}\selectfont}
papersize: a4
bibliography: child/references/references.bib
nocite: | 
  @*
csl: child/references/apa-6th-edition.csl
---

```{r echo = FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  message = FALSE,
  warning = FALSE,
  knitr.kable.NA = ''
)
```


```{r}
library(kableExtra)
library(dplyr)
library(tidyr)
library(MASS)
library(pROC)
library(caret)
library(knitr)
```


# Introduction {#sec:intro}


```{r, child = "child/01_introduction.Rmd", echo = FALSE}

```

# Related Literature


```{r, child = "child/02_rrl.Rmd", echo = FALSE}

```

# Methodology

```{r, child = "child/03_methodology.Rmd", echo = FALSE}

```

# Result

```{r, child = "child/04_result.Rmd"}

```

# Conclusion

```{r, child = "child/05_conclusion.Rmd"}

```

\clearpage
\vspace*{1cm}
\phantomsection

# References {-}

\vspace*{0.8cm}

::: {#refs}
:::
  
\clearpage
\vspace*{1cm}

\phantomsection 
\appendix


# Appendix {-}

\vspace*{0.8cm}

\setcounter{section}{0}
\renewcommand{\thesection}{A}

\setcounter{subsection}{0}
\renewcommand{\thesubsection}{A.\arabic{subsection}}

\numberwithin{table}{subsection}
\numberwithin{figure}{subsection}

```{r, child = "child/06_appendix.Rmd"}
```